Compositional Verification of Concurrent C Programs with Search Structure Templates

### 2023-11-23 concurrency/CPP 2024/paper.tex
resize table
first try to modify traverse spec

%\usepackage{txfonts}
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption
\usepackage{uri}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{mathpartir}
\usepackage{semantic}
\usepackage{graphicx}
\usepackage{cases}
\usepackage{hyperref}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{parcolumns}
%\usepackage{iris}
%\usepackage{lstlangcoq}
\usepackage[edges]{forest}
\renewcommand{\lstlistingname}{Figure}

@@

%\wm{The fonts are still off here---for instance, p, n, and n' are mathematical variables, not references to the C code. Fix if we have time.}
The \lstinline{traverse} function can then be specified as follows:
%\than{Try to avoid using specific name of $\mathrm{tree\_rep}$ for public condition, I suggested to use $\mathrm{Node}$ for generic purpose.} 
\begin{mathpar}
	% spec of inRange
	{\color{blue}
		\fforall \  m. \left\langle \texttt{pn} \mapsto (p, n) \ \ast \ 
		\nodeboxrep(n)  \ \ast \ \mathsf{R}(n) \ \big| \ \treerep\ (m) \
		\right\rangle
	}\vspace{-6pt}\\
	\vspace{-6pt}\texttt{traverse(pn, k)} \\
	{\color{blue}
		\left\langle \mathit{res}. \ \exists \  n', v.
		\begin{array}{l} \texttt{pn} \mapsto \left(n', n'\right) \ \ast \\ \nodeboxrep\left(n'\right) \ \ast\ \texttt{k} \in \mathsf{range}\ \ast \ 
			\\ 
			\mathsf{if} \ \mathit{res} \  \ \mathsf{then} \  \\ \ \ \ \mathsf{node\_contents}(n', \cdot, \mathsf{range}) \ 
			\\ \mathsf{else} \\ \ \ \ \mathsf{node\_contents}\left(n', (\texttt{k}, v), \mathsf{range}\right)
		\end{array}
		\ \Bigg | \treerep\ (m) \
		\right\rangle
	}
\end{mathpar}
The local precondition of \lstinline{traverse} includes both the node handle $\nodeboxrep(n)$ and its contents $\mathsf{R}(n)$, indicating that a thread should already hold $n$'s lock before calling \lstinline{traverse}. The output of \lstinline{traverse} is a new node $n'$ in the data structure such that the key \texttt{k} falls within the range of $n'$. The local postcondition then includes the handle of the new node $\nodeboxrep(n')$, its contents $\mathsf{R}(n')$, and a Boolean variable $\mathit{res}$ indicating whether \lstinline{traverse} found an empty node or a node with key \lstinline{k}.

@@

\begin{table}[h]
	\begin{center}
		\begin{tabular}{l | l || l | l } 
			\hline
			\textbf{Lock coupling} & \textbf{LoC} & \textbf{Give up} & \textbf{LoC} \\
			\hline
			Specfications & 190 & Specfications & 138 \\
			\texttt{findNext} proof 	& 27  & \texttt{findNext} proof & 27 \\
			\texttt{insertOp} proof 	& 32  & \texttt{insertOp} proof & 31 \\
			& & \texttt{inRange} proof & 30 \\
			\texttt{traverse} proof	& 360  & \texttt{traverse} proof & 336 \\ 
			\texttt{insert} proof 	   		& 450  & \texttt{insert} proof & 284 \\
			\texttt{lookup} proof 	   	& 235  &  \texttt{lookup} proof & 304 \\ 
			\texttt{pushdown\_left} proof 	   	& 282 & & \\	
			\texttt{delete} proof 		& 182	 &   & 	 \\
			supporting proofs    & 	1909	 & supporting proofs 	& 1876 \\
			\hline
			\textbf{Total} & \textbf{3667} & & \textbf{3026}  
		\end{tabular}
		\caption{Size of Coq definitions and proofs, by topic.}
		\label{table_coq}
	\end{center}
\end{table}

@@

Compared to prior template proofs in Iris and GRASShopper, our proof development is significantly larger (although direct comparison is difficult, since GRASShopper is an automatic prover). The biggest structural difference is that we need to directly state loop invariants for our imperative code, while the Iris proofs handle recursion with a simpler Löb-induction approach. Coming up with the loop invariants was the hardest part of the proofs. Our proofs also deal with the details of real C code: proving absence of integer overflows, accurately representing struct-and-pointer-based data structures instead of functional-programming-style structured data, etc. In exchange for this extra effort, the code we verify constitutes complete C programs that can be compiled and executed.
The full development of our mechanization effort is available online at \href{https://zenodo.org/record/8337004}{\color{blue}{https://zenodo.org/record/8337004}}.

>>>>

%\usepackage{txfonts}
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
%% http://ctan.org/pkg/subcaption
\usepackage{uri}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{mathpartir}
\usepackage{semantic}
\usepackage{graphicx}
\usepackage{cases}
\usepackage{hyperref}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{parcolumns}
\usepackage{scalerel}
%\usepackage{lstlangcoq}
\usepackage[edges]{forest}
\renewcommand{\lstlistingname}{Figure}

@@

%\wm{The fonts are still off here---for instance, p, n, and n' are mathematical variables, not references to the C code. Fix if we have time.}
The \lstinline{traverse} function can then be specified as follows:
%\than{Try to avoid using specific name of $\mathrm{tree\_rep}$ for public condition, I suggested to use $\mathrm{Node}$ for generic purpose.} 
\begin{mathpar}
{ \color{blue}
	\fforall \  m. \ \scaleleftright[1.50ex]{\langle}
	{\begin{array}{l}
			\texttt{pn} \mapsto (p, n) \ \ast \ 
			\nodeboxrep(n)  \ast \ \mathsf{R}(n) 
		\end{array}  
		\ \Bigg \vert  \treerep\ (m)
	}  
	{\rangle}
}
\vspace{-6pt}\\
\vspace{-6pt}\texttt{traverse(pn, k)} \\
{ \color{blue}
\scaleleftright[1.50ex]{\langle}
{\begin{array}{l}
		\mathit{res}. \ \exists \  n', v, \mathsf{range}. \\ 
		\ \ \ \ \  \texttt{pn} \mapsto \left(n', n'\right) \ast  \nodeboxrep\left(n'\right) \ast\ \texttt{k} \in \mathsf{range}\ \ast \ 
		\\ 
		\ \ \ \ \ \mathsf{if} \ \mathit{res}  \ \mathsf{then}  \ \mathsf{node\_contents}(n', \cdot, \mathsf{range}) \ 
		\\ 
		\ \ \ \ \ \ \ \ \ \ \ \ \ \ \mathsf{else} \ \mathsf{node\_contents}\left(n', (\texttt{k}, v), \mathsf{range}\right)
\end{array}  
\Bigg \vert \treerep\ (m)
}  
{\rangle}
}
\end{mathpar}

The local precondition of \lstinline{traverse} includes both the node handle $\nodeboxrep(n)$ and its contents $\mathsf{R}(n)$, indicating that a thread should already hold $n$'s lock before calling \lstinline{traverse}. The output of \lstinline{traverse} is a new node $n'$ in the data structure such that the key \texttt{k} falls within the range of $n'$. The local postcondition then includes the handle of the new node $\nodeboxrep(n')$, its contents $\mathsf{R}(n')$, and a Boolean variable $\mathit{res}$ indicating whether \lstinline{traverse} found an empty node or a node with key \lstinline{k}.

@@

\begin{table}[h]
	\centering
	\resizebox{\columnwidth}{!}{
		\begin{tabular}{l | l || l | l} 
			\hline
			\textbf{Lock coupling} & \textbf{LoC} & \textbf{Give up} & \textbf{LoC} \\
			\hline
			Specifications & 190 & Specifications & 138 \\
			\texttt{findNext} proof & 27 & \texttt{findNext} proof & 27 \\
			\texttt{insertOp} proof & 32 & \texttt{insertOp} proof & 31 \\
			& & \texttt{inRange} proof & 30 \\
			\texttt{traverse} proof & 360 & \texttt{traverse} proof & 336 \\ 
			\texttt{insert} proof & 450 & \texttt{insert} proof & 284 \\
			\texttt{lookup} proof & 235 & \texttt{lookup} proof & 304 \\ 
			\texttt{pushdown\_left} proof & 282 & & \\	
			\texttt{delete} proof & 182 & & \\
			Supporting proofs & 1909 & Supporting proofs & 1876 \\
			\hline
			\textbf{Total} & \textbf{3667} & & \textbf{3026}  
		\end{tabular}
	}
	\caption{Size of Coq definitions and proofs, by topic.}
	\label{table_coq}
\end{table}

@@

Compared to prior template proofs in Iris and GRASShopper, our proof development is significantly larger (although direct comparison is difficult, since GRASShopper is an automatic prover). The biggest structural difference is that we need to directly state loop invariants for our imperative code, while the Iris proofs handle recursion with a simpler Löb-induction approach. Coming up with the loop invariants was the hardest part of the proofs. Our proofs also deal with the details of real C code: proving absence of integer overflows, accurately representing struct-and-pointer-based data structures instead of functional-programming-style structured data, etc. In exchange for this extra effort, the code we verify constitutes complete C programs that can be compiled and executed.
The full development of our mechanization effort is available online at \href{https://zenodo.org/record/8337004}{https://zenodo.org/record/8337004}.
